---
title: "p8105_hw3_cl4469"
author: "Chen Liang"
date: "2023-10-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(ggplot2)
library(p8105.datasets)
```

```{r}
data("instacart")
```

```{r}
instacart %>% select(product_id) %>% distinct %>% count()

instacart %>% select(user_id, order_id) %>% distinct %>% count 
  
instacart %>% select(user_id) %>% distinct %>% count

```

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_obs=n()) |>
  arrange(desc(n_obs))

max(pull(instacart,aisle_id))
```
This is a table showing the number of items ordered from aisle. There are `r max(pull(instacart,aisle_id))` aisles.And the most items ordered from aisles are fresh vegetables and fresh fruits.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle=fct_reorder(aisle, n)) |>
  ggplot(aes(x=aisle,y=n))+
  geom_point()+
  labs(
    title="Number of items ordered in each aisle",
    x="Aisle",
    y="Number of items (n)",
    caption = "Data from the instacart package")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text = element_text(size = 7.5) )
```
Here is a plot that shows the number of items ordered in each aisle. And aisles are ordered by ascending number of items.

```{r}
rank_3 <- instacart |> 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) |>
  group_by(aisle,product_name)|>
  summarise(n=n()) |>
  mutate(product_ranking=min_rank(desc(n))) |>
  filter(product_ranking < 2)

knitr::kable(rank_3, caption = "Most Popular Item in Each of The Aisles")
```
This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_hour= mean(order_hour_of_day)) |>
  spread(key = order_dow, value = mean_hour) |>
  knitr::kable(caption = "Mean Hour of The Day 2 Items Are Ordered ", digits = 2)
```
This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. The result is a 2x7 table. There are 7 variables as column names ranging from 0 to 6, the row are the specific day in a week. The cell values are means of the hours of each day in a week. From this table, we can see Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

## Problem 2

```{r}
data("brfss_smart2010")
```

#### Data cleaning:
```{r}
brfss_df=
  brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic=="Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```
For this step in data cleaning, I organize responses as a factor taking levels ordered from “Poor” to “Excellent, and include only response from Excellent to poor.

```{r}
brfss_df |> 
  filter(year==2002 | year==2010) |>
  group_by(year,locationabbr,locationdesc) |>
  summarize(n=n()) |>
  count(locationabbr,name = "locations") %>% 
    filter(locations >= 7) %>% 
  pivot_wider(
    names_from = year,
    values_from = locations
  )
```
Comment: The table shows states observed equal or more than 7 times in both 2002 and 2010.There are 6 states have observed equal or more than 7 times in 2002, which are "CT" "FL" "MA" "NC" "NJ" "PA". However, in 2010, there are 14 states have observed equal or more than 7 times, which are "CA" "CO" "FL" "MA" "MD" "NC" "NE" "NJ" "NY" "OH" "PA" "SC" "TX" "WA".

Produce a plot where each state’s line depicts how the average “Excellent” response value changes over the years.
```{r}
brfss_df |> 
  filter(response == "Excellent") |>
  group_by(year,locationabbr) |>
  summarize(mean_data_value=mean(data_value,na.rm = TRUE)) |>
  ggplot(aes(x = year, y = mean_data_value, group = locationabbr,color = locationabbr)) + 
  geom_line()+ 
  labs(
    y = "Average of Data Value",
    title = "Spaghetti Plot: Average Value Over Time Within a State"
  ) +
  theme_minimal()
```
Comment:The spagetti plot shows the mean data values for excellent overall health in each state from 2022 to 2010. The average values for the majority of states appear to fluctuate within a relatively stable range, typically between 15 and 30. However, the data trends exhibit significant variations among the different states."

Now, we make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_df |>
  filter(year %in% c(2006,2010),
         locationabbr == "NY") |>
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_grid(.~year) +
  theme(axis.text.x = element_text()) +
  labs(
    title = "Density Plots for 2006 and 2010",
    x = "Response",
    y = "Density",
    caption = "Data from BRFSS Dataset"
  )
```


```{r}
brfss_df |>
  filter(year %in% c(2006,2010),
         locationabbr == "NY") |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot(aes(fill = response)) +
  facet_grid(.~year) +
  theme(axis.text.x = element_text()) +
  labs(
    title = "Box Plots for 2006 and 2010",
    x = "Response",
    y = "Data_value",
    caption = "Data from BRFSS Dataset"
  )
```
Comment:The density plots and boxplots shows the data values by responses (2006 vs 2010).In both 2006 and 2010, the data values increase overall from “poor” to “good” and “very good”, and the “good” and “very good” categories drew the highest number of responses in both 2006 and 2010. 

## Problem 3
Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

First, we load and tidy nhanes_accel data.
```{r}
accel_df <- read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    min1:min1440,
    names_to = "min",
    names_prefix = "min",
    values_to = "mims"
    )
```

And then, we load and tidy nhanes accelorometer data.
```{r}
covar_df <- read_csv("data/nhanes_covar.csv",skip = 4) |>
  janitor::clean_names() |>
  mutate(sex=ifelse(sex==1,"Male","Female"),
         sex = factor(sex, levels = c("Male", "Female")),
         education = case_match(education,
                                1 ~ "Less than high school",
                                2 ~ "High school equivalent",
                                3 ~ "More than high school"),
         education = factor(education, 
                            levels = c("Less than high school", "High school equivalent", "More than high school")))
```

Next, we merge the tidied accel_df and covar_df data 
```{r}
nhanes_df=
  inner_join(accel_df,covar_df)|>
  drop_na(sex,age,bmi,education) |>
  filter(age >= 21)
```

Description: The final combined dataframe named "nhanes_df" has `r nrow(nhanes_df)` observations and `r ncol(nhanes_df)` variables. This date set includes all initially observed variables and excludes participants younger than 21 years of age and those with missing demographic data.Additionally, we organized the education as a factor taking levels ordered from "Less than high school" to "More than high school" by using fct_relevel(), and the sex as a factor taking from "male" and "female".


Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

Create a table of the age distributions for men and women in each education category
```{r}
nhanes_df |>
  group_by(education,sex) |>
  summarize(n=n()) |>
  pivot_wider(
    names_from = sex,
    values_from = n) |>
    knitr::kable()
```

Comment: From this table and boxplot, we can derived that of the three levels of education, both males and females have the highest percentage with a "high school equivalent", compared to "Less than high school" and "More than high school". And there is no significant difference between number of male and female in education level of “Less than high school” and “More than high school”; however, the number of male in “High school equivalent” is significantly larger than that of women.



Create a visualization of the age distributions for men and women in each education category
```{r}
nhanes_df |> 
  ggplot(aes(x = education, y = age, color = sex)) + 
  geom_boxplot() +
  labs(title = "Age distribution for men and women", 
        x = "Education level",
        y = "Age")
```

Comment:The boxplot shows the min, max, and median ages of males and females for each level of educatiom. For each three of categories, ages are between 20 and 80 years. For “Less than High school education” and "high school equivalent education", we can see females have a higher median age than males. For the “More than high school education”,females have a lower median age than males.


Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences.
```{r}
nhanes_df |>
  group_by(seqn,age,sex,education) |>
  summarize(total_activity=sum(mims)) |>
  ggplot(aes(x=age,y=total_activity,color=sex))+
  geom_point(alpha = .7) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ education) +
  labs(title="Total activities against age for males and females",
       x="Age",
       y="Total activities")
```
Comment:For “less than high school education”, both male and female's total activity tends to decrease as age increases.And Total activity is generally higher in female than in male until about age 50. There is a peak around age 60 in female’s tend line but soon decline after that. For participants with a "high school equivalent" and "more than high school education", total activity of female is generally higher than male across all ages.


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.
```{r}
nhanes_df |>
  mutate(min = as.numeric(min)) |>
  ggplot(aes(x=min,y=mims, color=sex))+
  geom_point(alpha=0.2)+
  geom_smooth(se=F)+
  facet_grid(. ~ education)+
  labs(title="24-Hour activity time courses",
       x="Minute of a day",
       y="Mean activity",
       caption = "Data from nhanes dataset")+
  theme(legend.position = "bottom")
```

Comment:
The 24-hour activity time course plot is divided into three panels. Across all education levels, males consistently exhibit higher or equivalent activity levels compared to females. Notably, in the 'Less than High School' and 'High School Equivalent' education categories, female activity levels appear slightly lower than their male counterparts, particularly in the early time segments.However, within the 'More Than High School' education category, the data reveals the most pronounced peaks in activity. Specifically, males exhibit higher activity levels in the morning, while females are more active during the evening peak. The difference between male and female activity is most striking during the latter part of the day."

